GCE UROP Summary

### 2021-06-10 summaries/oct-2020/oct-2020.tex
Added .tex files

For the purposes of this summary, we will take the luminosity of the GCE to be \SI{6.37}{\erg \per\second}, which is consistent with the luminosity used in ref. \cite{fermilab}.

\subsection{Observations}
The \textit{Fermi} Space Telescope has observed two features which we explore in this summary. We model the sensitivity of the telescope as a step function, with the telescope detecting no pulsars of luminosity below $L_{t} = \SI{e34}{\erg\per\second}$ and all the pulsars above it.

@@

@@

Of these, the log-normal luminosity function is more physically relevant, but the power law is simpler. Both have been studied in detail by the papers cited, but the observations from \textit{Fermi} have not yet been applied to a log-normal luminosity function, so it is useful to compare the two in this light.

@@

\begin{figure}[h]
    \label{contour-overlay-power-law}
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/contour-overlay.png}
    \caption{Above is plotted the number of MSPs needed to reproduce the GCE under a power law luminosity with an exponential cutoff. In red is plotted the configurations required to match \textit{Fermi}'s observation of $R=1/4$, and in green their observation that $N_f = 47$. }
\end{figure}

The intersection point between these two families represents the only power law configuration that simultaneously realizes both of \textit{Fermi}'s observations, and it corresponds to $\si{3.3e6}$ MSPs in the GCE which is unrealistic. The region for $R < 1/4$ and $N_f < 47$, which is allowed by observation, appears beneath both curves.

@@

\begin{figure}[h]
    \label{contour-overlay-extra-power-law}
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/contour-overlay-extra.png}
    \caption{Above is plotted the number of MSPs needed to reproduce the GCE under a power law luminosity with an exponential cutoff. In red curves are plotted different values of $R$, with yellow representing higher values, while different values of $N_f$ are plotted in green, with brighter green representing higher $N_f$. The Fermilab results $R=1/4$ and $N_f=47$ are bolded.}
\end{figure}

@@

\begin{figure}[h]
    \label{contour-vary-alpha}
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/contour-vary-alpha-exp.png}
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/contour-vary-alpha-hard.png}
    \caption{Above is plotted the curves representing $R=1/4$ and $N_f=47$ for various $\alpha$ from 1.22 to 2.5. Higher $\alpha$ correspond to brighter colors. The blue curve represents the intersection points of these families for each $\alpha$. An exponential cutoff is used in the top plot, and a hard cutoff for the bottom.}
\end{figure}

@@

\begin{figure}[h]
    \label{contour-overlay-log-normal}
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/log-normal/contour-overlay.png}
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/log-normal/contour-overlay-extra.png}
    \caption{The number of MSPs needed to reproduce the GCE under a log-normal luminosity. On the top plot, the configurations required to match \textit{Fermi}'s observation of $R=1/4$ are shown in red, and in green their observation that $N_f = 47$. The plot is expanded for the lower plot to different values of $R$ and $N_f$, with higher $R$ and $N_f$ represented by brighter colors.}
\end{figure}

@@

\begin{thebibliography}{99}
    \bibitem{fermilab}Y. Zhong, S. McDermott, I. Cholis, P. Fox. ``A New Mask for An Old Suspect: Testing the Sensitivity of the Galactic Center Excess to the Point Source Mask.'' arXiv:1911.12369v (2019).

    \bibitem{log-normal}D. Hooper and T. Linden. ``The Gamma-Ray Pulsar Population of
    Globular Clusters: Implications for the
    GeV Excess.'' \textit{JCAP} (2016).

>>>>

For the purposes of this summary, we will take the luminosity of the GCE to be \SI{6.37e36}{\erg \per\second}, which is consistent with the luminosity used in ref. \cite{fermilab}.

\subsection{Observations}
The \textit{Fermi} Space Telescope has observed two features which we explore in this summary.

@@

We model the sensitivity of the telescope as a step function, with the telescope detecting no pulsars of luminosity below $L_{t} = \SI{e34}{\erg\per\second}$ and all the pulsars above it. Future parts of this study will test the sensitivity of our results on the value of $L_t$ and try more realistic sensitivity functions, such as that given by equation 2.34 of ref. \cite{ploeg}.

@@

Of these, the log-normal luminosity function is more physically relevant, but the power law is simpler. Both have been studied in detail by refs \cite{fermilab, log-normal}, but the observations from \textit{Fermi} have not yet been applied to a log-normal luminosity function, so it is useful to compare the two in this light.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/step-sensitivity/contour-overlay.png}
    \caption{Above is plotted the number of MSPs needed to reproduce the GCE under a power law luminosity with an exponential cutoff. In red is plotted the configurations required to match \textit{Fermi}'s observation of $R=1/4$, and in green their observation that $N_f = 47$. The circle represents the values of $L_{max}$ and $L_{min}$ used in ref. \cite{fermilab}.}
    \label{contour-overlay-power-law}
\end{figure}

The intersection point between these two families represents the only power law configuration that simultaneously realizes both of \textit{Fermi}'s observations, and it corresponds to $\SI{3.3e6}{}$ MSPs in the GCE which is unrealistic. The region for $R < 1/4$ and $N_f < 47$, which is allowed by observation, appears beneath both curves.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/step-sensitivity/contour-overlay-extra.png}
    \caption{Above is plotted the number of MSPs needed to reproduce the GCE under a power law luminosity with an exponential cutoff. In red curves are plotted different values of $R$, with yellow representing higher values, while different values of $N_f$ are plotted in green, with brighter green representing higher $N_f$. The \textit{Fermi} results $R=1/4$ and $N_f=47$ are bolded. The circle represents the values of $L_{max}$ and $L_{min}$ used in ref. \cite{fermilab}.}
    \label{contour-overlay-extra-power-law}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/step-sensitivity/contour-vary-alpha-exp.png}
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/power-law/step-sensitivity/contour-vary-alpha-hard.png}
    \caption{Above is plotted the curves representing $R=1/4$ and $N_f=47$ for various $\alpha$ from 1.22 to 2.5. Higher $\alpha$ correspond to brighter colors. The blue curve represents the intersection points of these families for each $\alpha$. An exponential cutoff is used in the top plot, and a hard cutoff for the bottom.}
    \label{contour-vary-alpha}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/log-normal/contour-overlay.png}
    \includegraphics[width=0.6\textwidth]{../../luminosity-models/log-normal/contour-overlay-extra.png}
    \caption{The number of MSPs needed to reproduce the GCE under a log-normal luminosity. On the top plot, the configurations required to match \textit{Fermi}'s observation of $R=1/4$ are shown in red, and in green their observation that $N_f = 47$. The plot is expanded for the lower plot to different values of $R$ and $N_f$, with higher $R$ and $N_f$ represented by brighter colors. The purple point represents the population configuration calculated by ref. \cite{log-normal} for globular clusters; the purple lines represent 1-$\sigma$ error bars around this point; and the black point is the configuration requiring the least pulsars within one standard deviation of the $\sigma$ calculated in ref. \cite{log-normal} while remaining within the observational constraints.}
    \label{contour-overlay-log-normal}
\end{figure}

@@

\begin{thebibliography}{99}
    \bibitem{fermilab} Y. Zhong, S. McDermott, I. Cholis, P. Fox. ``A New Mask for An Old Suspect: Testing the Sensitivity of the Galactic Center Excess to the Point Source Mask.'' arXiv:1911.12369v (2019).

    \bibitem{log-normal} D. Hooper and T. Linden. ``The Gamma-Ray Pulsar Population of
    Globular Clusters: Implications for the GeV Excess.'' \textit{JCAP} (2016).

    \bibitem{ploeg} H. Ploeg, C. Gordon, R. Crocker, O. Macias. ``Comparing the Galactic Bulge and Galactic Disk Millisecond Pulsars'' arXiv:2008.10821v2 (2020).
